### 2016-07-20 doc/publications/plos16/plos16.tex
Update discussion and table, remove some latin, update 20 2016 template

% Template for PLoS
% Version 3.1 February 2015
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig." instead of "Figure".
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,a4paper]{article}  %was \documentclass[10pt,letterpaper]{article}

\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

@@

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

@@

% rotating package for sideways tables
\usepackage{rotating}

@@

\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%hv
{\raggedright\footnotesize #1}}
\reversemarginpar
\setlength{\marginparwidth}{5cm}

@@

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{PLOS-submission}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

@@

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{Cursed Forest - A random forest implementation for `big' and `wide' data}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Aidan O'Brien\textsuperscript{1},
Piotr Szul\textsuperscript{2},
Stephanie Li\textsuperscript{3},
James Doecke\textsuperscript{3},
Nick Ellis\textsuperscript{4},
Robert Dunne\textsuperscript{5}, and
Denis C. Bauer\textsuperscript{1,*}
%, with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\bf{1} Health \& Biosecurity, CSIRO, Sydney, NSW, Australia
\\
\bf{2} Data61, CSIRO, Brisbane, QLD, Australia
\\
\bf{3} Health \& Biosecurity, CSIRO, Brisbane, QLD, Australia
\\
\bf{4} Oceans \& Atmosphere, CSIRO, Brisbane, QLD, Australia
\\
\bf{5} Data61, CSIRO, Sydney, NSW, Australia
\\
\bigskip

@@

\section*{Notes}
\begin{itemize}
\item what are you guys editing this with?  what about some carriage
  returns or line feeds or something!
\item Are we going to use the TCGA data in this paper?
\item are we going to run sparkML or Cursed forest on the extended
  simulation data \cite[]{Genuer.et.al.2010}? Has Aidan done this already?
\item I dont think I can get R to handle chromosme 1 etc. May have to run CF on chromosme 22, see table
  \ref{table:smaller_data}
\item check Piotr's simulation
\item run an argument as follows:
  \begin{itemize}
  \item variable  selection may succeed where accurate prediction does not
  \item consider  \cite{Segal.2004} examples
  \item  perhaps we argue that in the case where the relationship of $X$ and $y$ takes a particular form we can recover
    the reltionship for very large $p$p
  \end{itemize}
\item \cite{Tuv.et.al.2009} feature selection via adding permuted variables. Do we want to consider this?
\item chek out varSelRF
\end{itemize}

\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida
pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend
volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim
ultrices, ultrices elit pulvinar, volutpat risus.

@@

\subsection*{VariantSpark.}

@@

\section*{Random Forests converges on wide data}

@@

\subsection*{CursedForest processes genome wide variant data}
\label{comp}
Spark ML vs CursedForest vs Kmeans vs LR on small and full version (where applicable) of 1000G (performance and
time/mem)

% In this section we compare Apache's higher level machine learning library, Spark ML, with MLLib, the standard
% low-level library. From the results, we see that times are similar.  There are many practical benefits for using Spark
% ML with VariantSpark, with a selection of these benefits derived from the `pipeline' system Spark ML uses.  One
% benefit is that hyperparameter optimisation is built in. Although possible in

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Performance comparison between the different Machine Learning algorithms.}}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
%\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
\bf{Genome \%}  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
\hline
8\% (chr1) & CursedForest & XX & 1hr 58min 23sec & 4GB \\ \hline
          & RF (Spark ML) & XX & 8hr 29min 07sec & 8GB \\ \hline
          &  RF (R) & &  &\\ \hline
          &  LR & &  &\\ \hline
XX\% & CursedForest & &  &\\ \hline
          & RF(Spark ML) & &  &\\ \hline
          &  RF(R) & &  &\\ \hline
          &  LR & &  &\\ \hline

100\% & CursedForest & 0.96 & 7hr 14min 40sec & 8GB \\ \hline
          & RF (Spark ML) & - & - & - \\ \hline
          &  RF (R) & &  &\\ \hline
          &  LR & - & - & - \\ \hline
          & kmeans (Spark ML) & 0.82 & 30hr 44min 00sec & 24GB \\ \hline
\end{tabular}
\begin{flushleft} 
  Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus
  tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
%\end{adjustwidth}
\end{table}

\begin{table}
\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{|p{4.2cm}|l|l|l|l|l|l|l|}
\hline
\bf{Genome }  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
\hline
 chr 22 - phase1 (2.1GB)  &  RF (R)\footnote{10 nodes, 500 trees, dpas03} 
                                   &  0.456 & 2.89 & 18GB \\ \hline
\end{tabular}
\medskip
\begin{flushleft} 
\caption{some R results on smaller data sets.}
\label{table:smaller_data}
\end{flushleft}
\end{center}
%\end{sidewaystable}
\end{minipage}
\end{table}

% \begin
% \begin{table}[!ht]
% \caption{
% {\bf Performance comparison between the different Machine Learning algorithms.}}
% \begin{tabular}{|p{4.2cm}|l|l|l|l|l|l|l|}
% \hline
% \bf{Genome }  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
% \hline
%  chr 22 - phase1 (2.1GB)  &  RF (R) &  0.456 & 2.89 & 18GB \\ \hline
% \end{tabular}
% \begin{flushleft} 
% \end{flushleft}
% \label{table1}
% %\end{adjustwidth}
% \end{table}

@@

\subsection*{James' dataset}
Comparing CursedForest with feature selection + linear model (performance and time/mem)

@@

\subsection*{S1 Fig}
\label{S1_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S2 Fig}
\label{S2_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

>>>>

% Template for PLoS
% Version 3.3 June 2016
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

@@

@@

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

@@

@@

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

@@

\usepackage{multirow}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{Cursed Forest - A random forest implementation for `big' and `wide' data} % Please use "title case" (capitalize all terms in the title except conjunctions, prepositions, and articles).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Aidan O'Brien\textsuperscript{1},
Piotr Szul\textsuperscript{2},
Stephanie Li\textsuperscript{3},
James Doecke\textsuperscript{3},
Nick Ellis\textsuperscript{4},
Robert Dunne\textsuperscript{5}, and
Denis C. Bauer\textsuperscript{1,*}
%, with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\bf{1} Health \& Biosecurity, CSIRO, Sydney, NSW, Australia
\\
\bf{2} Data61, CSIRO, Brisbane, QLD, Australia
\\
\bf{3} Health \& Biosecurity, CSIRO, Brisbane, QLD, Australia
\\
\bf{4} Oceans \& Atmosphere, CSIRO, Brisbane, QLD, Australia
\\
\bf{5} Data61, CSIRO, Sydney, NSW, Australia
\\
\bigskip

@@

\section*{Notes}
\begin{itemize}
\item what are you guys editing this with?  what about some carriage
  returns or line feeds or something!
\item Are we going to use the TCGA data in this paper?
\item are we going to run sparkML or Cursed forest on the extended
  simulation data \cite[]{Genuer.et.al.2010}? Has Aidan done this already?
\item I dont think I can get R to handle chromosme 1 etc. May have to run CF on chromosme 22, see table
  \ref{table:smaller_data} -- On it, Aidan
\item check Piotr's simulation
\item run an argument as follows:
  \begin{itemize}
  \item variable  selection may succeed where accurate prediction does not
  \item consider  \cite{Segal.2004} examples
  \item  perhaps we argue that in the case where the relationship of $X$ and $y$ takes a particular form we can recover
    the reltionship for very large $p$p
  \end{itemize}
\item \cite{Tuv.et.al.2009} feature selection via adding permuted variables. Do we want to consider this?
\item chek out varSelRF
\end{itemize}

\section*{Abstract}
The abstract.

@@

\subsection*{VariantSpark.}
Traditionally, VariantSpark made use of Spark's machine learning algrorithms `Spark ML'. These algorithms require data in a `DataFrame',
similar to DataFrames in R.

We use the same pre-processing algorithms in CursedForests. One difference is that we

@@

\subsection*{Random Forests converges on wide data}

@@

\subsection*{CursedForest processes genome-wide variant data}
%\label{comp}
%Spark ML vs CursedForest vs Kmeans vs LR on small and full version (where applicable) of 1000G (performance and
%time/mem)

Traditionally, VariantSpark used the random forest algorithm from SparkML. While SparkML algorithms demonstrate
scalability when dealing with a large number of samples, this scalability quickly breaks down as we include more features.

One of the restrictions preventing SparkML algorithms from scaling to high-dimensionality data is the fact that the feature-set
for each individual is stored as a `Vector'. In Spark, Vectors are unable to be split and to process multiple vectors together,
the entirety of each vector must be loaded into memory on the same node. Therefore, as dimensionality increases, the
size of these Vectors must also increase, and subsequently this results in higher memory requirents.

On the other hand, CursedForest (CF) is specifically designed to handle wide ``cursed" data. It avoids the correlation between
memory and dimensionality by not using Vectors. Data is instead stored in a columnar form in a Parquet format.

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Performance comparison between the different machine learning algorithms.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
%\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
\bf{Genome \%}  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
\hline
\multirow{3}{*}{8\% (chr1: 6,450,364)} & CursedForest & XX & 1hr 58min 23sec & 4GB \\
& RF (Spark ML) & 0.94 & 8hr 29min 07sec & 8GB \\
& LR &  & 10hr 30min & 8GB\\ \hline

\multirow{3}{*}{24\% (chr1-3: 19,328,051)}  & CursedForest & &  &\\
& RF(Spark ML) & &  &\\
& LR & &  &\\ \hline

\multirow{4}{*}{100\% (chr1-22: 81,047,467)} & CursedForest & 0.96 & 7hr 14min 40sec & 8GB \\ 
& RF (Spark ML) & - & - & - \\
& LR & - & - & - \\ 
& kmeans (Spark ML) & 0.82 & 30hr 44min 00sec & 24GB \\ \hline
\end{tabular}
\begin{flushleft} 
Comparison of different various machine learning libraries on different subsets of variant data 
from the 1000 Genomes Project.

\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

\begin{table}
\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{|p{4.2cm}|l|l|l|l|l|l|l|}
\hline
\bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
\hline
RF (R)\footnote{10 nodes, 500 trees, dpas03} 
                                   &  0.456 & 2.89 & 18GB \\ \hline
CursedForest\footnote{something}
                                   & X & X & X \\ \hline
\end{tabular}
\medskip
\begin{flushleft} 
\caption{Comparion between Random Forests on R and CursedForests on a smaller
dataset. Chromosome 22 - phase1 (2.1GB) }
\label{table:smaller_data}
\end{flushleft}
\end{center}
%\end{sidewaystable}
\end{minipage}
\end{table}

% \begin
% \begin{table}[!ht]
% \caption{
% {\bf Performance comparison between the different Machine Learning algorithms.}}
% \begin{tabular}{|p{4.2cm}|l|l|l|l|l|l|l|}
% \hline
% \bf{Genome }  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
% \hline
%  chr 	 - phase1 (2.1GB)  &  RF (R) &  0.456 & 2.89 & 18GB \\ \hline
% \end{tabular}
% \begin{flushleft} 
% \end{flushleft}
% \label{table1}
% %\end{adjustwidth}
% \end{table}

@@

%\subsection*{James' dataset}
%Comparing CursedForest with feature selection + linear model (performance and time/mem)

@@
